\documentclass[a4paper,11pt]{book}

\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage[table]{xcolor}
\definecolor{lightblue}{HTML}{ADD8E6}
\definecolor{lightred}{HTML}{FF8D8D}
\definecolor{lightgreen}{HTML}{ADD816}
\usepackage[footnotesize,bf]{caption}
\usepackage{sidecap}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{SIunits} 
\usepackage{booktabs}
\usepackage[pdfborder={0 0 0}]{hyperref}

% trick to overwite rowcolor with columncolor
\makeatletter
\def\tmpp#1\@addtopreamble#2#3!{%
    \tmp#2!{#1}{#3}}
\def\tmp#1\CT@column@color\CT@row@color#2!#3#4{%
\def\@classz{#3\@addtopreamble{#1\CT@row@color\CT@column@color#2}#4}}
\expandafter\tmpp\@classz!
\makeatother

\usepackage{titlesec}
\titleformat{\chapter}[hang]{\bf\huge}{\thechapter}{0.5em}{}
\titlespacing{\chapter}{0pt}{0pt}{50pt}

\usepackage{natbib}
\setlength{\bibsep}{1pt}
\renewcommand{\bibfont}{\footnotesize}

\renewcommand*{\familydefault}{\sfdefault}

\hoffset -1in
\voffset -1in

\topmargin      10mm
\headheight     10mm
\headsep        10mm
\oddsidemargin  25mm
\textwidth     160mm
\textheight    242mm

%\title{\bf DNS/TMP User Guide}
\title{{\bf T}\textcolor{black!50}{Lab} User Guide}
\author{The T group}

\setcounter{tocdepth}{2}
\setlength{\parindent}{0mm}

\fancyhead[LE,LO]{\footnotesize\leftmark}
\fancyhead[RE,RO]{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\frontmatter
\pagestyle{empty}
\maketitle
\tableofcontents

\setlength{\parskip}{0.5\baselineskip}
\include{preface}

\mainmatter
\pagestyle{fancy}
\include{equations}
\include{bcs}
\include{code}
\include{grid}
\include{postprocessing}
\include{numerical}
\include{decomposition}
\include{scaling}

\backmatter
\bibliographystyle{plainnat}

%\bibliography{biblio}
\begin{thebibliography}{20}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Blaisdell et~al.(1996)Blaisdell, Spyropoulos, and Qin]{Blaisdell:1996}
G.~A. Blaisdell, E.~T. Spyropoulos, and J.~H. Qin.
\newblock The effect of the formulation of nonlinear terms on aliasing errors
  in spectral methods.
\newblock \emph{App. Num. Math.}, 21:\penalty0 207--219, 1996.

\bibitem[Carpenter and Kennedy(1994)]{Carpenter:1994}
M.~H. Carpenter and C.~A. Kennedy.
\newblock Fourth-order 2{N}-storage {R}unge-{K}utta schemes.
\newblock Technical Report TM-109112, NASA Langley Research Center, 1994.

\bibitem[Carpenter et~al.(1993)Carpenter, Gottlieb, and
  Abarbanel]{Carpenter:1993}
M.~H. Carpenter, D.~Gottlieb, and S.~Abarbanel.
\newblock The stability of numerical boundary treatments for compact high-order
  finite-difference schemes.
\newblock \emph{J. Comput. Phys.}, 108:\penalty0 272--295, 1993.

\bibitem[Erlebacher et~al.(1990)Erlebacher, Hussaini, Kreiss, and
  Sarkar]{Erlebacher:1990}
G.~Erlebacher, M.~Y. Hussaini, H.~O. Kreiss, and S.~Sarkar.
\newblock The analysis and simulation of compressible turbulence.
\newblock \emph{Theor. Comput. Fluid Dynamics}, 2:\penalty0 73--95, 1990.

\bibitem[Frigo and Johnson(2005)]{Frigo:2005}
M.~Frigo and S.~G. Johnson.
\newblock The design and implementation of {FFTW3}.
\newblock In \emph{Proceedings of the {IEEE}, vol. 93}, pages 216--231, 2005.

\bibitem[Higuera and Moser(1994)]{Higuera:1994}
F.~J. Higuera and R.~D. Moser.
\newblock Effect of chemical heat release in a temporally evolving mixing
  layer.
\newblock \emph{CTR Report}, pages 19--40, 1994.

\bibitem[Hu(1996)]{Hu:1996b}
F.~Q. Hu.
\newblock On absorbing boundary conditions for linearized {E}uler equations by
  a perfectly matched layer.
\newblock \emph{J. Comput. Phys.}, 129:\penalty0 201--219, 1996.

\bibitem[Hu et~al.(1996)Hu, Hussaini, and L.Manthey]{Hu:1996}
F.~Q. Hu, M.~Y. Hussaini, and J.~L.Manthey.
\newblock Low-dissipation and low-dispersion {R}unge-{K}utta schemes for
  computational acoustics.
\newblock \emph{J. Comput. Phys.}, 124:\penalty0 177--191, 1996.

\bibitem[Kravchenko and Moin(1997)]{Kravchenko:1997}
A.~G. Kravchenko and P.~Moin.
\newblock {On the effect of numerical errors in large-eddy simulations of
  turbulent flows}.
\newblock \emph{J. Comput. Phys.}, 131:\penalty0 310--322, 1997.

\bibitem[Lele(1992)]{Lele:1992}
S.~K. Lele.
\newblock {Compact finite difference schemes with spectral-like resolution}.
\newblock \emph{J. Comput. Phys.}, 103:\penalty0 16--42, 1992.

\bibitem[Lodato et~al.(2008)Lodato, Domingo, and Vervisch]{Lodato:2008}
G.~Lodato, P.~Domingo, and L.~Vervisch.
\newblock Three-dimensional boundary conditions for direct and large eddy
  simulation of compressible viscous flows.
\newblock \emph{J. Comput. Phys.}, 227:\penalty0 5105--5143, 2008.

\bibitem[Lomax et~al.(1998)Lomax, Pulliam, and Zingg]{Lomax:1998}
H.~Lomax, T.~H. Pulliam, and D.~W. Zingg.
\newblock \emph{Fundamentals of Computational Fluid Dynamics}.
\newblock Springer, 1998.

\bibitem[Mellado and Ansorge(2012)]{Mellado:2012}
J.~P. Mellado and C.~Ansorge.
\newblock Factorization of the {F}ourier transform of the pressure-{P}oisson
  equation using finite differences in colocated grids.
\newblock \emph{Z. Angew. Math. Mech.}, 92:\penalty0 380--392, 2012.

\bibitem[Shukla and Zhong(2005)]{Shukla:2005}
R.~K. Shukla and X.~Zhong.
\newblock Derivation of high-order compact finite difference schemes for
  non-uniform grid using polynomial interpolation.
\newblock \emph{J. Comput. Phys.}, 204:\penalty0 404--429, 2005.

\bibitem[Spalart et~al.(1991)Spalart, Moser, and Rogers]{Spalart:1991}
P.~R. Spalart, R.~D. Moser, and M.~M. Rogers.
\newblock Spectral methods for the {N}avier-{S}tokes equations with one
  infinite and two periodic directions.
\newblock \emph{J. Comput. Phys.}, 96:\penalty0 297--324, 1991.

\bibitem[Thompson(1987)]{Thompson:1987}
K.~W. Thompson.
\newblock Time-dependent boundary conditions for hyperbolic systems.
\newblock \emph{J. Comput. Phys.}, 68:\penalty0 1--24, 1987.

\bibitem[Thompson(1990)]{Thompson:1990}
K.~W. Thompson.
\newblock {Time-dependent boundary conditions for hyperbolic systems, II}.
\newblock \emph{J. Comput. Phys.}, 89:\penalty0 439--461, 1990.

\bibitem[Williams(1985)]{Williams:1985}
F.~A. Williams.
\newblock \emph{Combustion Theory}.
\newblock Addison Wesley, second edition, 1985.

\bibitem[Williamson(1980)]{Williamson:1980}
J.~H. Williamson.
\newblock {Low-storage Runge-Kutta schemes}.
\newblock \emph{J. Comput. Phys.}, 35:\penalty0 48--56, 1980.

\bibitem[Wilson et~al.(1998)Wilson, Demuren, and Carpenter]{Wilson:1998}
R.~V. Wilson, A.~O. Demuren, and M.~Carpenter.
\newblock Higher-order compact schemes for numerical simulation of
  incompressible flows.
\newblock Technical Report CR-1998-206922, NASA Langley Research Center, 1998.

\end{thebibliography}

\end{document}

