\chapter{Post-Processing Tools}\label{sec:postprocessing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Averages}

See file {\tt dns/tools/statistics/averages}.

Allows for conditional analysis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probability Density Functions}

See file {\tt dns/tools/statistics/pdfs}.

Allows for conditional analysis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conditional Analysis}

To be developed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Two-point Statistics}
\sloppy

See file {\tt dns/tools/statistics/spectra}. Based on package FFTW \cite{Frigo:2005}.

Given two scalar fields $\{a_{nm}:\,n=1,\ldots,N,\,m=1,\ldots,M\}$ and similarly
$b_{nm}$, we calculate the one-dimensional co-spectra
$\{E^x_0,\,E^x_1,\,\ldots,\,E^x_{N/2}\}$ and
$\{E^z_0,\,E^z_1,\,\ldots,\,E^z_{M/2}\}$ normalized such that
\begin{equation}
\langle ab\rangle = E^x_0+2\sum_1^{N/2-1}E^x_n+E^x_{N/2} = E^z_1+2\sum_0^{M/2-1}E^z_m+E^z_{M/2}
\end{equation}
The mean value is removed, such that the left-hand side is $\langle
a'b'\rangle$. The Nyquist frequency energy content $E^x_{N/2}$ and $E^z_{M/2}$
is not written to disk, only the $N/2$ values
$\{E^x_0,\,E^x_1,\,\ldots,\,E^x_{N/2-1}\}$ and the $M/2$ values
$\{E^z_0,\,E^z_1,\,\ldots,\,E^z_{M/2-1}\}$. When $a\equiv b$, then we obtain the
power spectral density.

The sum above can be interpreted as the trapezoidal-rule approximation to the
integral $(L/2\pi)\int_0^{\kappa_c}2E(\kappa)\mathrm{d}\kappa$, where
$\kappa_c=\pi/h$ is the Nyquist frequency, $\Delta \kappa=\kappa_c/(N/2)=2\pi/L$
is the uniform wavenumber spacing, $h=L/N$ is the uniform grid spacing and $L$
is the domain size. Hence, the physical spectral function at wavenumber $\kappa_n=
n\Delta \kappa$ (equivalently, wavelength $L/n$) is $2E_n/\Delta \kappa$.

Due to the relatively large size of the files, we split the calculations is the
auto-spectra and the cross-spectra. The corresponding files containing the
one-dimensional spectra along the direction $Ox$ are {\tt xsp} and {\tt xCsp},
respectively, and similarly along the direction $Oz$. The two-dimensional
co-spectra $E_{nm}$ can also be written to disk, though the additional memory
requirement can be a difficulty.

The one-dimensional cross-correlations $\{C^x_0,\,C^x_1,\,\ldots,\,C^x_{N-1}\}$
and $\{C^z_0,\,C^z_1,\,\ldots,\,C^z_{M-1}\}$ are normalized by
$a_\mathrm{rms}b_\mathrm{rms}$, so that $C^x_0 = C^z_0 =1$ when $a\equiv b$ and
we calculate the auto-correlations. The auto-correlations are even functions and
therefore only $\{C^x_0,\,C^x_1,\,\ldots,\,C^x_{N/2-1}\}$ and
$\{C^z_0,\,C^z_1,\,\ldots,\,C^z_{M/2-1}\}$ are written to disk (note that we
also dropped the last term $C^x_{N/2}$ and $C^z_{M/2}$.)

The corresponding files containing the one-dimensional cross-correlations along
the direction $Ox$ are {\tt xcr} and {\tt xCcr}, and similarly along the
direction $Oz$. The two-dimensional cross-correlation $C_{nm}$ can also be
written to disk, though the additional memory requirement can be a difficulty.

Both form a Fourier pair according to
\begin{equation*}
E_k=\frac{1}{N}\sum_0^{N-1}(a_\mathrm{rms}b_\mathrm{rms}C_n)\exp(-i\omega_kn)\;,\qquad
a_\mathrm{rms}b_\mathrm{rms}C_n=\sum_0^{N-1}E_k\exp(i\omega_kn)\;,
\end{equation*}
where $\{\omega_k=(2\pi/N)k:\, k = 0,\ldots,N-1\}$ is the scaled wavenumber and
$i=\sqrt{-1}$ is the imaginary unit. Therefore,
\begin{equation}
\frac{1}{N}\sum_0^{N-1}C_n=\frac{E_0}{a_\mathrm{rms}b_\mathrm{rms}} \;,
\end{equation}
relation that can be used to relate integral scales $\ell$ to the Fourier mode
$E_0$, as follows. First, for the auto-correlation function, we can re-write
\begin{equation}
\frac{1}{N/2}\sum_0^{N/2-1}C_n=\frac{E_0}{a_\mathrm{rms}^2}+\frac{1-C_{N/2}}{N} 
\end{equation}
because
\begin{equation*}
\frac{1}{N}\sum_0^{N-1}C_n = \frac{1}{N}\left(\sum_0^{N/2-1}C_n
+C_{N/2}+\sum_{N/2+1}^{N-1}C_n\right) = \frac{1}{N}\left(2\sum_0^{N/2-1}C_n
+C_{N/2}-1\right) \;,
\end{equation*}
since, from periodicity, $C_N=C_0=1$ and, from the symmetry of the
auto-correlation sequence, $\sum_{N/2+1}^{N}C_n=\sum_0^{N/2-1}C_n$. Therefore,
if we use a trapezoidal rule to define the integral length scale as
\begin{equation}
\ell=h\left(\frac{C_0+C_{N/2-1}}{2}+\sum_1^{N/2-2}C_n\right)\;,
\end{equation}
where $h=L/N$ is the grid spacing and $L$ is the domain size, we obtain
\begin{equation}
\ell=\frac{L}{2}\left(\frac{E_0}{a_\mathrm{rms}^2}-\frac{2C_{N/2}}{N}\right)
\simeq \frac{L}{2a_\mathrm{rms}^2}E_0\;.
\end{equation}
This result applies to both directions $Ox$ and $Oz$, providing relations
between $\ell^x$ and $E^x$, and $\ell^z$ and $E^z$. Each case needs to use the
corresponding domain size, $L^x$ and $L^z$.

These relations show that the integral length scales can be obtained directly
from the spectral information without the need to calculate the correlation
functions. However, the statistical convergence of those integral scales might
be too poor and alternative definitions might be more useful. Also, correlation
functions provide information about the degree of de-correlation achieved with a
particular domain size, and about the structural organization of the flow in
terms of different properties.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary of Budget Equations for Second-Order Moments}

\subsection{Reynolds Stresses}

Reynolds averages are indicated by a line and $u^{'}=u-\bar{u}$.  Favre averages
are used for quantities per unit mass and are indicated by a tilde,
e.g. $\tilde{u}=\overline{\rho u}/\overline{\rho}$ and $u^{''}=u-\tilde{u}$ and
$\widetilde{u^{''2}}=\overline{\rho u^{''2}}/\bar{\rho}$. In case of constant
density, Favre and Reynolds averages coincide.

The momentum equation written in non conservative form is
\begin{equation}
\rho \frac{\partial u_i}{\partial t} + \rho u_k \frac{\partial u_i}{\partial
  x_k} = -\frac{\partial p}{\partial x_i} + \frac{\partial \tau_{ik}}{\partial
  x_k} + b\,g_i - \epsilon_{imk} c_m\,\rho u_k
\end{equation}
where the non-dimensional numbers $Re$, $Fr$ and $Ro$ are included in the
corresponding tensors $\tau$, $g$ and $c$ for notational convenience.

Multiplying by $u^{''}_j$ and averaging
\begin{equation}
\label{eq:ap2}
\overline{\rho u^{''}_j \frac{\partial (\tilde{u}_i+u^{''}_i)}{\partial t}} +
\overline{\rho u^{''}_j (\tilde{u}_k+u^{''}_k)\frac{\partial (\tilde{u}_i+u^{''}_i)}{\partial x_k} } =
- \overline{u^{''}_j \frac{\partial p}{\partial x_i}} +
\overline{u^{''}_j \frac{\partial \tau_{ik}}{\partial x_k} }
+\overline{bu^{''}_j}\,g_i-\epsilon_{imk} c_m\overline{\rho u^{''}_j(\tilde{u}_k+u^{''}_k)}
\end{equation}
noting that $\overline{\rho u^{''}_j}=0$ we can simplify Eq. (\ref{eq:ap2}) to
\begin{equation}
\label{eq:ap3}
\overline{\rho u^{''}_j \frac{\partial u^{''}_i}{\partial t}} +
\overline{\rho u^{''}_j u^{''}_k }\frac{\partial \tilde{u}_i}{\partial x_k} +
\overline{\rho u^{''}_j \frac{\partial u^{''}_i}{\partial x_k} } \tilde{u}_k +
\overline{\rho u^{''}_k u^{''}_j \frac{\partial u^{''}_i}{\partial x_k} }  =
- \overline{u^{''}_j \frac{\partial p}{\partial x_i}} +
\overline{u^{''}_j \frac{\partial \tau_{ik}}{\partial x_k} }
+\overline{bu^{''}_j}\,g_i-\epsilon_{imk} c_m\overline{\rho u^{''}_ju^{''}_k}
\end{equation}
adding Eq. (\ref{eq:ap3}) with the indexes exchanged we get
\begin{equation}\label{eq:ap4}
  \begin{split}
    \overline{\rho \frac{\partial (u^{''}_i u^{''}_j)}{\partial t}} +
     \overline{\rho \frac{\partial (u^{''}_i u^{''}_j)}{\partial x_k}}
    \tilde{u}_k = & - (\overline{\rho u^{''}_k u^{''}_i} \frac{\partial 
      \tilde{u}_j}{\partial x_k}+\overline{\rho u^{''}_k u^{''}_j} 
    \frac{\partial \tilde{u}_i}{\partial x_k})
    - \overline{\rho u^{''}_k \frac{\partial (u^{''}_i u^{''}_j)}{\partial 
        x_k}} \\
    & - (\overline{u^{''}_j \frac{\partial p}{\partial x_i}+u^{''}_i 
      \frac{\partial p}{\partial x_j}}) + (\overline{u^{''}_j 
      \frac{\partial \tau_{ik}}{\partial x_k}+u^{''}_i \frac{\partial 
        \tau_{jk}}{\partial x_k}}) \\
    & + (\overline{bu^{''}_j}\,g_i+\overline{bu^{''}_i}\,g_j)\\
    & - (\epsilon_{imk} c_m\overline{\rho u^{''}_ju^{''}_k} +
         \epsilon_{jmk} c_m\overline{\rho u^{''}_iu^{''}_k} )
  \end{split}
\end{equation}

From the continuity equation 
\begin{equation}
  \frac{\partial \rho}{\partial t} + \frac{\partial (\rho u_k)}{\partial 
  x_k} = 0
\end{equation}
and multiplying by $u^{''}_i u^{''}_j$ and averaging
\begin{equation}
\label{eq:ap5}
  \overline{u^{''}_i u^{''}_j \frac{\partial \rho}{\partial t}} +
  \overline{u^{''}_i u^{''}_j \frac{\partial (\rho u^{''}_k)}{
      \partial x_k}} + \overline{u^{''}_i u^{''}_j \frac{\partial 
      (\rho \tilde{u}_k) }{\partial x_k} } = 0
\end{equation}
defining
\begin{equation}
  R_{ij} = \frac{\overline{\rho u^{''}_i u^{''}_j}}{\bar{\rho}}
\end{equation}
and adding Eqs. (\ref{eq:ap4}) and (\ref{eq:ap5}) we get
\begin{equation}
\begin{split}
  \frac{\partial (\bar{\rho} R_{ij})}{\partial t} +
  \frac{\partial (\bar{\rho} \tilde{u}_k R_{ij})}{\partial x_k} = 
  & -\bar{\rho}\biggl( R_{ik} \frac{\partial \tilde{u}_j}{\partial x_k} +
  R_{jk} \frac{\partial \tilde{u}_i}{\partial x_k} \biggr) -
  \biggl( \frac{\partial (\overline{\rho u^{''}_i u^{''}_j u^{''}_k}) }{
    \partial x_k} \biggr) \\
  & - \biggl( \overline{u^{''}_i \frac{\partial 
      p^{'}}{\partial x_j} + u^{''}_j \frac{\partial p^{'}}{\partial 
      x_i} } \biggr) + \biggl( \overline{u^{''}_i \frac{\partial 
      \tau^{'}_{jk}}{\partial x_k} + u^{''}_j \frac{\partial 
      \tau^{'}_{ik}}{\partial x_k}} \biggr) \\
  & - \biggl( \bar{u}^{''}_i 
  \frac{\partial \bar{p}}{\partial x_j} + \bar{u}^{''}_j \frac{\partial 
    \bar{p}}{ \partial x_i} \biggr) + \biggl( \bar{u}^{''}_i 
  \frac{\partial \bar{\tau}_{jk}}{\partial x_k} + \bar{u}^{''}_j 
  \frac{\partial \bar{\tau}_{ik}}{\partial x_k} \biggr) \\
  & + (\overline{bu^{''}_j}\,g_i+\overline{bu^{''}_i}\,g_j)
    - \bar{\rho}c_m(\epsilon_{imk} R_{jk} + \epsilon_{jmk} R_{ik} )
\end{split}
\end{equation}

The pressure and viscous terms can be decomposed as:
\begin{equation}
  \begin{split}
    \biggl( \overline{ u^{''}_i \frac{\partial 
        p^{'}}{\partial x_j} + u^{''}_j \frac{\partial p^{'}}{\partial 
        x_i} } \biggr) = & \biggl( \frac{\partial (\overline{u^{''}_i 
        p^{'}}) }{\partial x_j} + \frac{\partial (\overline{u^{''}_i 
        p^{'}})}{\partial x_i} \biggr) - \overline{p^{'}(
      \frac{\partial u^{''}_i}{\partial x_j}+\frac{\partial u^{''}_j}{
        \partial x_i})}\\
    = & \frac{\partial}{\partial x_k} \biggl(\overline{u^{''}_i 
        p^{'} \delta_{jk} + u^{''}_j p^{'} \delta_{ik}   }\biggr) 
      - \overline{p^{'}( \frac{\partial u^{''}_i}{\partial x_j}+
        \frac{\partial u^{''}_j}{\partial x_i})}\\
      \biggl( \overline{u^{''}_i \frac{\partial 
      \tau^{'}_{jk}}{\partial x_k} + u^{''}_j \frac{\partial 
      \tau^{'}_{ik}}{\partial x_k}} \biggr) = & \frac{\partial}{
    \partial x_k} \biggl( \overline{u^{''}_i \tau^{'}_{jk} +
    u^{''}_j \tau^{'}_{ik}} \biggr) - \biggl( \overline{\frac{
      \partial u^{''}_i }{\partial x_k} \tau^{'}_{jk} + \frac{
      \partial u^{''}_j }{\partial x_k} \tau^{'}_{ik}} \biggr) \\
  \end{split}
\end{equation}

Finally, the Reynolds stress equation reads:
\begin{equation}
  \frac{\partial R_{ij}}{\partial t}  = C_{ij} - F_{ij} + P_{ij}  + B_{ij} - \varepsilon_{ij} +
  \frac{1}{\bar{\rho}}\left(\Pi_{ij} - \frac{\partial T_{ijk}}{ \partial
    x_k}  + \Sigma_{ij} \right)
\end{equation}
where
\begin{eqnarray*}
  C_{ij} =&  - \tilde{u}_k \frac{\partial
    R_{ij}}{\partial x_k} & ,\text{advection} \\
  F_{ij} =&  \epsilon_{imk}c_m R_{jk} + \epsilon_{jmk}c_m R_{ik} &,
  \text{Coriolis redistribution}\\
  P_{ij} =&  - \biggl( R_{ik} \frac{\partial \tilde{u}_j}{
    \partial x_k} + R_{jk} \frac{\partial \tilde{u}_i}{
    \partial x_k}\biggr) & ,\text{turbulent shear production} \\
  B_{ij} =& \frac{1}{\bar{\rho}} \biggl(
  \overline{bu^{''}_j}\,g_i+\overline{bu^{''}_i}\,g_j \biggr) &,
  \text{turbulent buoyancy production}\\
  \varepsilon_{ij} =& \frac{1}{\bar{\rho}} \biggl( \overline{ 
    \tau^{'}_{jk}
    \frac{\partial u^{''}_i}{\partial x_k} + \tau^{'}_{ik}
    \frac{\partial u^{''}_j}{\partial x_k}}\biggr) & , \text{turbulent dissipation}\\
  T_{ijk} =& \overline{\rho u^{''}_i u^{''}_j u^{''}_k} +
  \overline{p^{'}u^{'}_i} \delta_{jk} + \overline{p^{'}u^{'}_j} 
  \delta_{ik} - ( \overline{\tau^{'}_{jk} u^{''}_i} + 
  \overline{\tau^{'}_{ik} u^{''}_j}) & , \text{turbulent transport} \\
  \Pi_{ij} =& \overline{p^{'}\biggl( \frac{\partial u^{''}_i}{
      \partial x_j}+ \frac{\partial u^{''}_j}{\partial x_i} \biggr)} & , \text{pressure strain} \\
  \Sigma_{ij} =&  \overline{u^{''}_i} \biggl( \frac{\partial \bar{\tau}_{jk}}{
    \partial x_k} - \frac{\partial \bar{p}}{\partial x_j} \biggr) +  
  \overline{u^{''}_j} \biggl( \frac{\partial \bar{\tau}_{ik}}{
    \partial x_k} - \frac{\partial \bar{p}}{\partial x_i} \biggr)& , \text{mean flux}\\
\end{eqnarray*}
Depending on symmetries, many of these terms are zero (within statistical
convergence) and substitutes for ensemble average can be used, like plane
averages of time averages.  Note that if $b\equiv \rho$, then $B_{ij}=0$. The
mean flux term is sometimes written as $\Sigma_{ij}=D_{ij}-G_{ij}$, the first
term grouping the mean viscous stress contributions and the last term the mean
pressure contributions. In cases of constant density, then
$\overline{u^{''}_j}=0$ and $\Sigma_{ij}=D_{ij}=G_{ij}=0$.

Contracting indices, the budget equation for the turbulent kinetic energy
$K=R_{ii}/2$ reads
\begin{equation}
  \frac{\partial K}{\partial t}  = C + P  + B - \varepsilon +
  \frac{1}{\bar{\rho}}\left(\Pi - \frac{\partial T_{k}}{ \partial
    x_k}  + \Sigma \right)
\end{equation}
Note that $F_{ii}=0$, always. If the flow is solenoidal, then $\Pi=0$.

\subsection{Energy Equation}

To be developed (before in terms of the pressure)

\subsection{Scalar Variance}

Similarly, the equation for the scalar r.m.s. is obtained from the scalar
conservation equations,

\begin{eqnarray*}
  \frac{\partial}{\partial t}(\rho \zeta) + 
  \frac{\partial}{\partial x_k}(\rho \zeta u_k) & = & 
  -\frac{\partial j_k}{\partial x_k}
  + w
\end{eqnarray*}
multiplying by $\zeta^{''}$ and averaging
\begin{equation}
  \label{eq:scalar1}
  \overline{\rho \zeta^{''} \frac{\partial \zeta^{''}}{\partial t}}
  + \overline{\rho \zeta^{''} u^{''}_k} \frac{\partial \tilde{\zeta}}{\partial x_k}
  + \overline{\rho \zeta^{''} \frac{\partial \zeta^{''}}{\partial x_k}} \tilde{u}_k
  + \overline{\rho \zeta^{''} u^{''}_k \frac{\partial \zeta^{''}}{\partial x_k}}
  = \\
  -\overline{\zeta^{''}}\frac{\partial \bar{j_k}}{\partial x_k}
  -\frac{\partial }{\partial x_k}\bigg(\overline{j_k' \zeta^{''}} \bigg)
  +\overline{j_k' \frac{\partial \zeta^{''}}{\partial x_k}}
  + \overline{w\zeta^{''}}
\end{equation}

Adding the mass conservation equation multiplied by $\frac{1}{2}\zeta^{''2}$
and averaging
\begin{multline}
  \label{eq:scalar2}
  \frac{\partial}{\partial t}(\overline{\rho \zeta^{''2}})
  + \frac{\partial }{\partial x_k}(\tilde{u}_k \overline{\rho \zeta^{''2}}) =\\
  -2 \overline{\rho \zeta^{''} u^{''}_k} \frac{\partial \tilde{\zeta}}{
    \partial x_k} - \frac{\partial}{\partial x_k}\bigg( 
  \overline{\rho \zeta^{''2} u^{''}_k} + 2\overline{j_k' \zeta^{''} }
  \bigg) +2 \overline{j_k'\frac{\partial \zeta^{''}}{\partial x_k}}
  -2\overline{\zeta^{''}}\frac{\partial \bar{j}_k}{\partial x_k}
  + 2 \overline{w\zeta^{''}}
\end{multline}

Defining
\begin{eqnarray}
  R_{\zeta\zeta} & = & \frac{\overline{\rho \zeta^{''2}}}{\bar{\rho}} \\
  R_{k \zeta} & = & \frac{\overline{\rho \zeta^{''} u^{''}_k}}{\bar{\rho}}
\end{eqnarray}
the previous transport equation can be further simplified to
\begin{equation}
  \label{eq:scalar3}
  \frac{\partial R_{\zeta \zeta}}{\partial t} + 
  \tilde{u}_k \frac{\partial R_{\zeta \zeta}}{\partial x_k}
  = P_{\zeta\zeta} - \chi +\frac{1}{\bar{\rho}} \left(
  - \frac{\partial T_{\zeta\zeta k}}{\partial x_k} + D_{\zeta\zeta} + Q_{\zeta\zeta}\right)
\end{equation}
where
\begin{eqnarray*}
  P_{\zeta \zeta} = & -2 R_{k \zeta} \frac{\partial \tilde{\zeta}}{\partial x_k}
  &,\text{turbulent production}\\
  \chi = & -\frac{2}{\bar{\rho}} \overline{j_k' \frac{\partial \zeta^{''}}{\partial x_k}}
  &,\text{turbulent dissipation}\\ 
  T_{\zeta\zeta k} = & \overline{\rho \zeta^{''2} u^{''}_k } +2
  \overline{j_k'\zeta^{''}}
  &,\text{turbulent transport}\\ 
  D_{\zeta\zeta} = & -2 \overline{\zeta^{''}} \frac{\partial \bar{j}_k}{\partial x_k} 
     &,\text{mean flux}\\
  Q_{\zeta\zeta} = & 2\overline{w\zeta^{''}}&,\text{source}
\end{eqnarray*}

